---
title: "Clustering Grocery Items"
author: "Mitul Shah"
date: "8/16/2017"
output: pdf_document
---


```{r}

## Loading the libraries
library(magrittr)
library(dplyr)

## Loading datasets
item_to_id <- read.csv("grocery/item_to_id.csv")
data <- read.csv("grocery/purchase_history.csv")

## Order the datasets 
item_to_id <- item_to_id %>% arrange(Item_id)
data <- data %>% arrange(user_id)

## Loading splitstackshape
library(splitstackshape)

## Creating dat
dat <- cSplit(data, "id", ",")

## Removing user id column
dat <- select(dat, -1)

## Creating grocery dataset
grocery <- data.frame(matrix(nrow = 39474, ncol = 48))

## Renaming columns to items
colnames(grocery) <- item_to_id$Item_id

## Create x as sequence from 1 to 48
x <- seq(1,48,1)

## Enter 1 if the transaction had that item and 0 if that item was not in the transaction
for(i in 1:nrow(dat)) {
  for(j in 1:ncol(dat)) {
    if( (dat[[i,j]] %in% x) == TRUE) {
      y <- dat[[i, j]]
      grocery[[i, y]] = 1
    } else {
      grocery[[i, y]] = 0
    }
  }
}

## Enter 1 if the transaction had that item and 0 if that item was not in the transaction
#for(i in 1:nrow(dat)) {
#  for(j in 1:ncol(dat)) {
#   if(is.na(dat[[i,j]]) == FALSE)  {
#      y <- dat[[i, j]]
#      grocery[[i, y]] = 1
#    } else {
#      grocery[[i, y]] = 0
#    } 
#  }
#}

grocery[grocery == 0] <- 1
grocery[is.na(grocery)] <- 0

## Renaming columns to Item name
colnames(grocery) <- item_to_id$Item_name


```


## Forming Clusters of Grocery Items

```{r}

## Transpose grocery data
grocery_data_to_cluster <- as.data.frame(t(grocery))

## Distance matrix
grocery.dist <- dist(grocery_data_to_cluster, method = "euclidean", diag = FALSE, upper = FALSE, p = 2)

## Hierarchical Clustering using Ward's method
grocery.hclust <- hclust(grocery.dist, method = "ward.D")

## Visualize the dendogram
plot(grocery.hclust, labels = item_to_id$Item_name, main='Dendogram')


```

Looking at the dendogram, grouping items into 12 clusters seems to be a good number. 

```{r}

## Forming 12 clusters
groups.12 <- cutree(grocery.hclust, 12)

## Looking at the items in all 12 clusters
sapply(unique(groups.12), function(g)item_to_id$Item_name[groups.12 == g])

## Comparing results with k-means forming 12 clusters
kmeans(grocery_data_to_cluster, 12, alg="Lloyd")[[1]]
```


The clusters formed by hierarchical clustering (Ward's method) seems to be better than the results shown by k-means clustering by looking at the names of the items in clusters. The results of k-means groups has 1 cluster with many items which doesn't look good to me. However, we might obtain better clusters by using k-mediods (using PAM) or using other linkage methods like single or complete linkage in hierarchical clustering. Looking at the above results, I would group the items by the results given by hierarchical clustering. 

## Finding customers who bought most items in her lifetime

```{r}

## Merge user id with grocery
data_with_users <- as.data.frame(cbind(data$user_id, grocery))

## Rename 1st column to user id
colnames(data_with_users)[1] <- "user_id"

## Data giving number of items bought by each customer in each transaction
data_to_find_customers_buying_most_items <- data_with_users %>% mutate(total_items_in_each_transaction = rowSums(.[2:49]))

## Grouping by customer
number_of_items_by_customer <- data_to_find_customers_buying_most_items %>% group_by(user_id) %>% summarise(total_items = sum(total_items_in_each_transaction))

## Finding maxiumum items bought by any customer
max(number_of_items_by_customer$total_items)

## The user id of the customer who bought most items 
filter(number_of_items_by_customer, total_items == 72)

```


The customer who bought the maximum items in her lifetime has the user id 269335. 


## Finding for each item, the customer who bought that product the most

```{r}

## Data giving how many times each user bought each item in all transactions
d <- data_with_users %>% group_by(user_id) %>% summarise_all(funs(sum))

## Creating data to show the customer who buys that item most number of times 
most_buying_customer_for_each_item <- data.frame(matrix(nrow = 48, ncol = 81))

## Renaming columns 
colnames(most_buying_customer_for_each_item) <- c("Item_name", paste0( "user_id_", 1:80))

## 1st column as the name of the item 
most_buying_customer_for_each_item$Item_name <- item_to_id$Item_name

for(i in 2:ncol(d)) {
  z <- max(d[,i])
  a <- filter(d, d[,i] == z)
  for(j in 1:nrow(a)) {
    most_buying_customer_for_each_item[[(i-1), (j+1)]] = a[[j, 1]]
  }
}

x <- c()

## All the users (they might be repeated)
for(i in 1:48) {
  for(j in 2:81) {
    y <- most_buying_customer_for_each_item[[i, j]]
    if (is.na(y) == TRUE) {
      x <- x
    } else {
      x <- c(x, y)
    }
  }
}

k <- c()

## Unique users 
for(i in 1:length(x)) {
  if(x[i] %in% k == FALSE) {
    k <- c(k, x[i])
  } else {
    k <- k
  }
}

## Print all users
k 
```

